### 2020-06-28 chapters/problem-statement.tex
EMA-like histogram
Method description done
problem statement

\textcolor{red}{INCOMPLETE}

@@

@@

    \item \textbf{Explainable alerts:} understand why each alert was raised
\end{itemize}

\section{Research Questions}
The goal of this Thesis is to develop a lightweight stream monitoring system that detects changes in features distribution between a reference period and a sliding window in real-time.

To that end, we devise a series of research questions to be answered by the end of this Thesis:

\begin{enumerate}
    \item \textbf{RQ1:}
    \item \textbf{RQ2:}
    \item \textbf{RQ3:}
\end{enumerate}

>>>>

@@

\section{The problem}
Many real-time stream monitoring systems are static once deployed in a production environment. Engineers of said systems configure them under the assumption that future data flowing through the system follows roughly the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. The problem at hand is determining when to reconfigure the system, whatever it may be (\textit{e.g.} a Machine Learning model), based on streaming analysis.

@@

    \item \textbf{Explainable alerts:} have some measure of divergence to understand why an alert was raised
\end{itemize}

\section{Proposal}
We devise a two-phased (batch and streaming phases) and two-windowed (reference and target sliding windows) method. We devise a sliding window approximate histogram aggregation that encodes data distributions and provide a method for incrementally maintaining it with constant time and space complexity for every growing volumes of data, ideal for the streaming scenario. Given a user-defined reference period and a stream of data to process, our method guarantees to monitor each event' fields and alert if they are considered divergent from their reference distribution. We provide a measure of how divergent each field (or feature) is. Based on the number of divergent features and the corresponding divergence measure, a system administrator has an easier time determining when to reconfigure the system.

\section{Research Questions}
The goal of this Thesis is to develop a lightweight stream monitoring system that detects changes in features distribution between a reference period and a sliding window in real-time. To that end, we devise a series of research questions to be answered by the end of this Thesis:

\begin{enumerate}[label=\textbf{(RQ\arabic*)}]
    \item What aggregation can we use to properly encode data distributions?
    \item How can we maintain said aggregation in a sliding window fashion in real-time with constant memory usage?
    \item How can we measure divergence between the reference and sliding window aggregations?
    \item Based on a divergence measure, when should we raise an alert?
\end{enumerate}

### 2020-06-28 chapters/validation.tex
problem statement

\section{Experiments with real data}

>>>>

\clearpage
\section{Experiments with real data}

Due to time constraints, we experimented with data from only one merchant. The merchant's entity will remain anonymous due to legal reasons. \textcolor{red}{merchancs use-case, how to define?}

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets from DATEX to DATEY. We used the older one as the reference dataset and the other one as the target one. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 182 numerical and categorical features. Both datasets were known to suffer a lot from concept drift. Additionally, the data scientists at Feedzai performed some data preprocessing and cleaning, removing some of the missing values, but only on the reference dataset, which further augmented the differences between the target and reference datasets. Because of these reasons, we expected a lot of divergence alarms.

TALK ABOUT DATASETS (size)

In our first analysis, for the batch phase, we used 100 Spark executors, each with 12G of memory. We also determined that one week worth of data for this merchant was on average equivalent to 37,803 events. Hence, we set the half-life $n_{1/2}=37,803$ and discard events whose contribution is 6\% or less, which is equivalent to using windows of data of four times the half-life (as detailed in Section \ref{sec:ema-hist}).
